---
title: "Crypto & Commodity"
author: "Jingsong, GAO"
date: "2021/10/10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dendextend)
library(dplyr)
library(data.table)
library(dtwclust)
library(factoextra)
library(ggdendro)
library(gplots)
library(ggplot2)
library(ggthemes)
library(ggthemr)
library(gridExtra)
library(tidyr)
library(tibble)
```

## 1. Preparation

Load Crypto price data.
```{r}
crypto = read.csv("../data/coin_metrics/PriceUSD.csv")
crypto$date = as.Date(crypto$date)
crypto = crypto[(nrow(crypto)-719):nrow(crypto),]
str(crypto)
```

Normalize all prices.
```{r}
for (i in 2:length(colnames(crypto))) {
  crypto[i] = crypto[i] / max(crypto[i])
}
row.names(crypto) <- crypto$date
crypto[1] <- NULL
head(crypto)
```

Transpose the data frame.
```{r}
crypto_ts = transpose(crypto)
row.names(crypto_ts) = colnames(crypto)
```


Load Commodity price data.
```{r}
commodity = read.csv("../data/commodity.csv")
commodity$Date = as.Date(commodity$Date)
commodity = commodity[(nrow(commodity)-719):nrow(commodity),]
str(commodity)
```

Normalize all prices.
```{r}
for (i in 2:length(colnames(commodity))) {
  commodity[i] = commodity[i] / max(commodity[i])
}
row.names(commodity) <- commodity$date
commodity[1] <- NULL
head(commodity)
```

Transpose the data frame.
```{r}
commodity_ts = transpose(commodity)
row.names(commodity_ts) = colnames(commodity)
```


Load Amazon products.
```{r}
amazon = read.csv("../data/amazon/NEW_eqts_2y.txt", header = F, sep = " ")
amazon_label = read.csv("../data/amazon/NEW_label_2y.txt", header = F)$V1
```


Normalize all prices.
```{r}
for (i in 1:length(amazon[,1])) {
  amazon[i,] = amazon[i,] / max(amazon[i,])
}
```


## 2. KMeans clustering
We train the KMeans clustering models with $k = 4:10$. Then we evaluate the models.
```{r}
km1 = tsclust(amazon, type = "partitional", k = 4L:10L, 
              distance = "dtw_basic", centroid = "pam", 
              seed = 511L)
names(km1) = paste0("k_", 4L:10L)
sapply(km1, cvi, type = "internal")
```
The majority result from these evaluation factors suggest that $k=5$ is the best number of clusters.

```{r}
kmeans = tsclust(amazon, type = "partitional", k = 5L, 
              distance = "dtw_basic", centroid = "pam", 
              seed = 511L)
```

```{r}
gpc = plot(kmeans, plot=FALSE)
gpc + 
  ylim(0,1) +
  labs(x="Time", y="Normalized Price")
ggsave("picture/r_2y_kmeans5.png")
```

Show cluster table.
```{r}
cluster_result = data.frame(label=tstrsplit(amazon_label, split=" ")[[1]], cluster=kmeans@cluster)
result_table = aggregate(rep(1,nrow(cluster_result)), by=list(cluster=cluster_result$cluster ,label=cluster_result$label), FUN=length)

result_table
```


## 3. Cluster Crypto data
Use the trained model to predict all crypto data.
```{r}
crypto_labels = predict(kmeans, newdata = crypto_ts)
crypto_labels
```

```{r}
p.crypto = crypto %>%
  mutate(index = 1:720) %>%
  gather(key=type_col, value=value, -index) %>%
  ggplot(aes(x=index, y = value)) +
  geom_line() +
  facet_wrap(~type_col, ncol = 1, strip.position="right") + 
  guides(color="none") + 
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.text.y = element_text(angle = 0),
        strip.background = element_blank(),
        panel.spacing.y = unit(0,"line")) +
  labs(x="Time", y="Price")
ggsave("picture/crypto.png", p.crypto, height=6, width=5)
```


## 4. Cluster Commodity data
Use the trained model to predict all commodity data.
```{r}
commodity_label = predict(kmeans, newdata = commodity_ts)
print(commodity_label)
commodity_label = data.frame(clst = commodity_label, type_col=colnames(commodity))
```

```{r}
p.comm = commodity %>%
  mutate(index = 1:720) %>%
  gather(key=type_col, value=value, -index) %>%
  full_join(., commodity_label, by = "type_col") %>%
  ggplot(aes(x=index, y = value, colour=factor(clst))) +
  geom_line() + 
  scale_colour_brewer(palette = "Set1") +
  facet_wrap(~type_col, ncol = 1, strip.position="right") +
  guides(color="none") + 
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.text.y = element_text(angle = 0),
        strip.background = element_blank(),
        panel.spacing.y = unit(0,"line")) +
  labs(x="Time", y="Price")
ggsave("picture/commodity.png", p.comm, height=5, width=5)
```

## 5. Hierarchical clustering
Then we use hierarchical clustering.
```{r}
ts_amazon = amazon %>%
  transpose(.) %>%
  rename_all(funs(c(paste(cluster_result$label,.,sep="_"))))

hc1 = tsclust(t(ts_amazon), type = "h", k = 5L, 
              distance = "dtw_basic", centroid = shape_extraction,
              seed = 511L)
```

```{r}
hclus <- cutree(hc1, k = 2) %>%
  as.data.frame(.) %>%
  rename(.,cluster_group = .) %>%
  rownames_to_column("type_col")
```


```{r}

#p1 = hcdata %>%
#  ggdendrogram(., rotate=TRUE, leaf_labels=FALSE)

den = as.dendrogram(hc1) %>%
  set("branches_k_color", k=5)
label = sapply(labels(den), function(x) unlist(strsplit(x, "_"))[1])
bars = sapply(label, function(x) {
  if (x == "gpu") return("red")
  if (x == "cpu") return("gold")
  if (x == "drive") return("blue")
})
png(file="picture/amazon_dend.png", width=36, height=9, units="in", res=600)
par(mar = c(10,2,1,1))
plot(den)
colored_bars(colors = bars, dend = den, sort_by_labels_order = FALSE, rowLabels = "Product Name")
dev.off()
```


```{r}
hcdata = dendro_data(hc1)
names_order = hcdata$labels$label
p2 = ts_amazon %>%
  mutate(index = 1:720) %>%
  gather(key=type_col, value=value, -index) %>%
  full_join(., hclus, by = "type_col") %>%
  mutate(type_col = factor(type_col, levels = rev(as.character(names_order)))) %>% 
  ggplot(aes(x=index, y = value, colour = factor(cluster_group))) +
  geom_line() +
  facet_wrap(~type_col, ncol = 1, strip.position="right") + 
  scale_colour_brewer(palette = "Set1") +
  guides(color="none") + 
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.text.y = element_text(angle = 0),
        strip.background = element_blank(),
        panel.spacing.y = unit(0,"line"))
ggsave("picture/amazon_ts.png", p2, height=48, width=2)
```

Compute distances.
```{r}
dist.eu = dist(t(ts_amazon), method = "euclidean")
dist.man = dist(t(ts_amazon), method = "manhattan")
hc = tsclust(t(ts_amazon), type = "h", k = 5L, 
              distance = "dtw_basic", centroid = shape_extraction,
              seed = 511L)
dist.dtw = dist(hc@distmat)

cosine.sim = function(X) {
    X = as.matrix(X)
    sim = X / sqrt(rowSums(X * X))
    return(dist(sim %*% t(sim)))
}
dist.cos = cosine.sim(t(ts_amazon))
```


```{r}
gdist.eu = fviz_dist(dist.eu,
                     order=F,
                     gradient = list(low = "red", mid = "white", high = "blue")) +
    theme_economist_white() +
    theme(
        title = element_text(size = 20),
        axis.text.y = element_text(size = 6),
        axis.text.x = element_blank(),
        panel.grid.minor.x = element_blank()
    ) +
    labs(title = "Visualization of Euclidean Distances")
ggsave(gdist.eu, filename = "picture/r_dist_eu.png",
       device = "png", height = 20, width = 20, dpi = 300)

gdist.dtw = fviz_dist(dist.dtw, 
                      order=F, 
                      gradient = list(low = "red", mid = "white", high = "blue")) +
    theme_economist_white() +
    theme(
        title = element_text(size = 20),
        axis.text.y = element_text(size = 6),
        axis.text.x = element_blank(),
        panel.grid.minor.x = element_blank()
    ) +
    labs(title = "Visualization of Dynamic Time Warping Distances")
ggsave(gdist.dtw, filename = "picture/r_dist_dtw.png",
       device = "png", height = 20, width = 20, dpi = 300)

gdist.cos = fviz_dist(dist.cos,
                      order=F,
                      gradient = list(low = "red", mid = "white", high = "blue")) +
    theme_economist_white() +
    theme(
        title = element_text(size = 20),
        axis.text.y = element_text(size = 6),
        axis.text.x = element_blank(),
        panel.grid.minor.x = element_blank()
    ) +
    labs(title = "Visualization of Cosine Similarity")
ggsave(gdist.cos, filename = "picture/r_dist_cos.png",
       device = "png", height = 20, width = 20, dpi = 300)
```



